---
title: "Kaggle_other trials"
author: "Shiya Wang"
date: "11/09/2021"
output: html_document
---

# Data transformation tried 

Transform transit with "train" or "subway" as 1, others as 0
```{r}
premium_transit = c("train","subway")
pattern2 = paste(premium_transit,collapse="|")

df$transit = ifelse(grepl(pattern2,df$transit),1,0)

```

Transform zipcode, matching zipcode with median market listing price of each zipcode (another dataset on zipcode housing price)
```{r}
library(data.table)
library(stringr)

# load market price dataset
price_zip = fread('C:\\2. Course\\5200 AA Frameworks_Lala Fri\\Kaggle project\\RDC_Inventory_Core_Metrics_Zip_History.csv',select = c('month_date_yyyymm','postal_code','median_listing_price_per_square_foot'))

# for airbnb dataset: clean the zipcode (e.g. "NY 10024" to "10024", "  10024" to "10024")
df$zipcode = ifelse(startsWith(df$zipcode,'NY'),str_sub(df$zipcode, -5,-1),str_sub(df$zipcode,1,5))
df$zipcode = ifelse(startsWith(df$zipcode,' '),gsub(' ','',df$zipcode),df$zipcode)

# for market price dataset: calculate the average monthly median listing price in 2020
price_zip = price_zip %>%
  filter(startsWith(month_date_yyyymm,'2020'))
price_zip$median_listing_price_per_square_foot[price_zip$median_listing_price_per_square_foot<=200] = NA

x = tapply(price_zip$median_listing_price_per_square_foot,price_zip$postal_code,mean)
y = data.frame(x)
colnames(y) = 'median_price'
price_NYzip = cbind(zipcode = rownames(y),y)
rownames(price_NYzip) = 1:nrow(price_NYzip)
head(price_NYzip)

#merge airbnb dataset with market price dataset
df = merge(df,price_NYzip,by='zipcode',all.x=TRUE)
df$median_price[is.na(df$median_price)] = median(df$median_price, na.rm = TRUE)
df = subset(df,select = -c(zipcode))
price_zip[price_zip$postal_code=='10024',]

```

# indicator selection (exclude highly correlated variables)

Check highly correlated variable
```{r}
colnames(df)
# [4] "host_listings_count" [5] "host_total_listings_count" [39] "calculated_host_listings_count"[40] "calculated_host_listings_count_entire_homes" [41] "calculated_host_listings_count_private_rooms"
df01 = df[,c(4,5,39:41,15)]
# [20] "maximum_minimum_nights" [21] "minimum_maximum_nights" [22] "maximum_maximum_nights" [23] "minimum_nights_avg_ntm" [24] "maximum_nights_avg_ntm"  
df02 = df[,c(20:24,15)]
# [26] "availability_30" [27] "availability_60"[28] "availability_90" [29] "availability_365" 
df03 = df[,c(26:29,15)]
#[30] "number_of_reviews"[31] "number_of_reviews_ltm"  
df04 = df[,c(30:31,15)]

ggpairs(df01)
#"host_listings_count" [5] "host_total_listings_count" =1; [39] "calculated_host_listings_count"[40] "calculated_host_listings_count_entire_homes**"=0.97
ggpairs(df02)
#[21] "minimum_maximum_nights" [22] "maximum_maximum_nights" [24] "maximum_nights_avg_ntm" = 1
ggpairs(df03)
# "availability_30**" "availability_60**" = 0.94;  "availability_60" "availability_90" = 0.97
ggpairs(df04)
```

Exclude highly correlated variable
```{r}
#cor = 1
df = subset(df,select = -c(host_total_listings_count))
df = subset(df,select = -c(minimum_maximum_nights,maximum_maximum_nights))
ncol(df)
#cor > 0.9
#df = subset(df,select = -c(calculated_host_listings_count,availability_60,availability_90))

```

threshold 0.02
```{r}
#df = subset(df,select = -c(host_is_superhost,maximum_minimum_nights,minimum_nights_avg_ntm,maximum_nights_avg_ntm,review_scores_checkin,reviews_per_month))
```


# Other models tried

## SVM

Linear SVM
```{r}
library(e1071)


svmLinear = svm(price~.,data=train_dummy ,kernel='sigmoid') #,type='eps-regression', scale=FALSE)
pred_train = predict(svmLinear,data=train_dummy)
rmse_train = sqrt(mean((pred_train - train_dummy$price)^2)); rmse_train
pred = predict(svmLinear,newdata=test_dummy)
rmse = sqrt(mean((pred - test_dummy$price)^2)); rmse
```


```{r}
library(e1071)

svmRadial = svm(price~.,kernel='radial',data=train) #,kernel='linear',type='eps-regression'
pred_train = predict(svmRadial,data=train)
rmse_train = sqrt(mean((pred_train - train$price)^2)); rmse_train
pred = predict(svmRadial,data=test)
rmse = sqrt(mean((pred - test$price)^2)); rmse
```


Normalize numerical variables
```{r}

# normalize <- function(x, na.rm = TRUE) {
#     return((x- min(x)) /(max(x)-min(x)))
# }
# y = c(1,3,5)
# normalize(y)
# 
# data = as.data.frame(apply(df$host_listings_count, FUN = function(x) normalize(x)))
# 
# apply(X = train[,c(2:3,8:26)], MARGIN = 2, FUN = function(x) normalize(x))


pp = preProcess(train[,-10],method="range")
colnames(train)

train_i = predict(pp,train[,-10])
test_i = predict(pp,test[,-10])

train_nmz = as.data.frame(cbind(price = train$price,train_i))
test_nmz = as.data.frame(cbind(price = test$price,test_i))

train_nmz

```


Dummy coding categorical variables
```{r}
colnames(train_nmz)
dmy = dummyVars("~.",data = train_nmz[,-1])
train_i2 = predict(dmy,newdata = train_nmz[,-1])
train_dummy = as.data.frame(cbind(price = train_nmz$price,train_i2))

dmy = dummyVars("~.",data = test_nmz[,-1])
test_i2 = predict(dmy,newdata = test_nmz[,-1])
test_dummy = as.data.frame(cbind(price = test_nmz$price,test_i2))

```

Radial SVM
```{r}
library(e1071)


svmLinear = svm(price~.,data=train_dummy ,kernel='radial') #,type='eps-regression', scale=FALSE)
pred_train = predict(svmLinear,data=train_dummy)
rmse_train = sqrt(mean((pred_train - train_dummy$price)^2)); rmse_train
pred = predict(svmLinear,newdata=test_dummy)
rmse = sqrt(mean((pred - test_dummy$price)^2)); rmse
```


```{r}
pred_test_ <- predict(svmLinear,newdata=test_dummy);
rmse = sqrt(mean((pred_test_ - test_dummy$price)^2)); rmse
head(pred_test)
#head(pred_train)
nrow(pred_train)

mean(train_dummy$price);mean(test_dummy$price);
mean(pred_train);mean(pred)
```



## Tree

Default tree
```{r}
library(rpart); library(rpart.plot)
tree = rpart(price~.,data = train, method = 'anova')
pred_train = predict(tree)
rmse_train_tree = sqrt(mean((pred_train - train$price)^2)); rmse_train_tree
pred = predict(tree, newdata = test)
rmse_tree = sqrt(mean((pred - test$price)^2)); rmse_tree
```

Tuned tree
```{r}
tuneGrid = expand.grid(cp = seq(0,0.1,0.001))
trControl = trainControl(method = 'cv',number = 5)
set.seed(1000)
tree_cv = train(price~.,
               data = train,
               method = 'rpart',
               trControl = trControl, 
               tuneGrid = tuneGrid)
```


```{r}
library(rpart); library(rpart.plot)
tuned_tree = rpart(price~.,data = train, method = 'anova',cp=tree_cv$bestTune)
pred_train = predict(tuned_tree)
rmse_train_tuned_tree = sqrt(mean((pred_train - train$price)^2)); rmse_train_tuned_tree
pred = predict(tuned_tree, newdata = test)
rmse_tuned_tree = sqrt(mean((pred - test$price)^2)); rmse_tuned_tree
```

## Ranger

ranger
```{r}
colnames(train)
library(ranger)
set.seed(1031)
forest_ranger = ranger(price~.,
                       data = train, 
                       mtry = 25,
                       num.trees = 500,
                       )
pred_train = predict(forest_ranger, data = train, num.trees = 500)
rmse_train_forest_ranger = sqrt(mean((pred_train$predictions - train$price)^2)); rmse_train_forest_ranger
pred = predict(forest_ranger, data = test, num.trees = 500)
rmse_forest_ranger = sqrt(mean((pred$predictions - test$price)^2)); rmse_forest_ranger
```

tuned ranger
```{r}
library(ranger)
trControl=trainControl(method='cv',number=5)
tuneGrid=expand.grid(mtry=c(20,25,30,35),splitrule="variance",min.node.size=200)
#c('variance','extratrees','maxstat')
set.seed(1031)
cvModel = train(price~.,
                data = train, 
                method='ranger',
                num.trees = 500,
                trControl=trControl,
                tuneGrid=tuneGrid)
cvModel$bestTune

pred_train = predict(forest_ranger, data = train, num.trees = 1000)
rmse_train_forest_ranger = sqrt(mean((pred_train$predictions - train$price)^2)); rmse_train_forest_ranger
pred = predict(forest_ranger, data = test, num.trees = 1000)
rmse_forest_ranger = sqrt(mean((pred$predictions - test$price)^2)); rmse_forest_ranger
```

## gbm

tuned gbm
```{r}
library(caret)
set.seed(1031)
trControl = trainControl(method="cv",number=5)
tuneGrid = expand.grid(n.trees = 500, 
                       interaction.depth = c(1,2,3),
                       shrinkage = c(0.001,0.01,0.1),
                       n.minobsinnode=c(5,10,15))
garbage = capture.output(cvModel <- train(price~.,
                                          data=train,
                                          method="gbm",
                                          trControl=trControl, 
                                          tuneGrid=tuneGrid))
set.seed(1031)
cvboost = gbm(price~.,
              data=train,
              distribution="gaussian",
              n.trees=500,
              interaction.depth=cvModel$bestTune$interaction.depth,
              shrinkage=cvModel$bestTune$shrinkage,
              n.minobsinnode = cvModel$bestTune$n.minobsinnode)

pred_train = predict(cvboost, n.trees=500)
rmse_train_cv_boost = sqrt(mean((pred_train - train$price)^2)); rmse_train_cv_boost
```
